---
title: "E1 and E2 Data Analysis"
output: html_notebook
---

# Comparison of the move distributions between E1 and E2 and across Goal Types

```{r}
library(tidyverse)

e1 <- read_csv("data/e1/final_tracker.csv")
e2 <- read_csv("data/e2/final_tracker.csv")

e1_architects <- e1 %>%
  filter(role == "architect") %>%
  mutate(experiment = "E1")

e2 <- e2 %>%
  mutate(experiment = "E2")

combined <- bind_rows(e1_architects, e2)

combined <- combined %>%
  mutate(
    move_utility = case_when(
      move_utility == 1 ~ "useful",
      move_utility == 0 ~ "inconsequential",
      move_utility == -1 ~ "harmful"
    ),
    move_type = case_when(
      move_utility == "useful" ~ "Useful",
      move_utility == "inconsequential" ~ "Inconsequential",
      move_utility == "harmful" ~ "Harmful"
    )
  )

# total number of different move types
move_summary <- combined %>%
  group_by(experiment, move_utility, goal_type) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = move_utility, values_from = count, values_fill = 0) %>%
  rename(
    harmful = `harmful`,
    inconsequential = `inconsequential`,
    useful = `useful`
  ) %>%
  mutate(total_moves = harmful + inconsequential + useful)

# average distribution of moves per participant, grouped by goal type and experiment
avg_moves_by_goal <- combined %>%
  group_by(experiment, ID, anonID, goal_type, move_type) %>%
  summarise(count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = move_type, values_from = count, values_fill = 0) %>%
  mutate(total_moves = Useful + Inconsequential + Harmful)


# COUNTING BY PROPORTION and Summary stats
prop_moves <- combined %>%
  group_by(ID, goal_type, experiment) %>%
  count(move_type, name = "move_count") %>%
  mutate(total_moves = sum(move_count),
         prop = move_count / total_moves) %>%
  ungroup()

prop_wide <- prop_moves %>%
  select(-move_count) %>%
  pivot_wider(
    names_from = move_type,
    values_from = prop,
    values_fill = 0
  )

print(prop_wide)

prop_moves_summary <- prop_wide %>%
  group_by(goal_type, experiment) %>%
  summarise(
    mean_Useful = mean(Useful),
    sd_Useful = sd(Useful),
    mean_Inconsequential = mean(Inconsequential),
    sd_Inconsequential = sd(Inconsequential),
    mean_Harmful = mean(Harmful),
    sd_Harmful = sd(Harmful),
  )
print(prop_moves_summary)

# summary statistics: mean and SD of move types across goal types and experiments
avg_moves_summary <- avg_moves_by_goal %>%
  group_by(goal_type, experiment) %>%
  summarise(
    mean_Useful = mean(Useful),
    sd_Useful = sd(Useful),
    mean_Inconsequential = mean(Inconsequential),
    sd_Inconsequential = sd(Inconsequential),
    mean_Harmful = mean(Harmful),
    sd_Harmful = sd(Harmful),
    mean_total_moves = mean(total_moves),
    sd_total_moves = sd(total_moves)
  )
print(avg_moves_summary)

```

##### Visualization of move distributions

```{r}
library(ggplot2)

ggplot(avg_moves_summary, aes(x = goal_type, y = mean_Useful, fill = experiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) + 
  geom_errorbar(aes(ymin = mean_Useful - sd_Useful, ymax = mean_Useful + sd_Useful),
  position = position_dodge(width = 0.9),
  width = 0.2) +
  labs(title = "Players' Avg Number of Useful Moves Per Goal Type by Experiment",
       x = "Goal Type", y = "Average Useful Moves Per Player", fill = "Experiment" ) +
  theme_minimal()

ggplot(avg_moves_summary, aes(x = goal_type, y = mean_Harmful, fill = experiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = mean_Harmful - sd_Harmful, ymax = mean_Harmful + sd_Harmful),
  position = position_dodge(width = 0.9),
  width = 0.2) +
  labs(title = "Players' Avg Number of Harmful Moves Per Goal Type by Experiment",
       x = "Goal Type", y = "Average Harmful Moves Per Player", fill = "Experiment") +
  theme_minimal()

ggplot(avg_moves_summary, aes(x = goal_type, y = mean_Inconsequential, fill = experiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = mean_Inconsequential - sd_Inconsequential, ymax = mean_Inconsequential + sd_Inconsequential),
  position = position_dodge(width = 0.9),
  width = 0.2) +
  labs(title = "Players' Avg Number of Inconsequential Moves Per Goal Type by Experiment",
       x = "Goal Type", y = "Average Inconsequential Moves Per Player", fill = "Experiment") +
  theme_minimal()

# BY PROPORTION

library(dplyr)
library(ggplot2)

# 2) Bar‐chart with error bars
ggplot(prop_summary, aes(x = move_type, y = mean_prop, fill = experiment)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, alpha = 0.8) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                position = position_dodge(width = 0.8),
                width = 0.2) +
  labs(
    title = "Mean Proportion of Move Types by Experiment",
    x     = "Move Type",
    y     = "Mean Proportion",
    fill  = "Experiment"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )


```

##### Statistical Tests Comparing Move Types Across Experiments

E2 players made significantly more useful moves than E1 players across all goal-types. 

For the "move" goal type, E2 players also made significantly more inconsequential moves than E1 players. 



# load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(emmeans)
```

# import data

```{r}
e1 <- read_csv("data/e1/final_tracker.csv")
e2 <- read_csv("data/e2/final_tracker.csv")%>%
  mutate(experiment = "E2")
```


# E1 vs E2 move analysis

```{r}
combined <- bind_rows(e1 %>%
  filter(role == "architect") %>%
  mutate(experiment = "E1"), e2)%>%
  mutate(
    move_type = case_when(
      move_utility == 1 ~ "Useful",
      move_utility == 0 ~ "Inconsequential",
      move_utility == -1 ~ "Harmful"
    ),
  ID = ifelse(experiment == "E2", anonID, ID))
```
### model
```{r}
# COUNTING BY PROPORTION and Summary stats
prop_moves <- combined %>%
  count(experiment, ID, goal_type, move_type, name = "move_count") %>%
  group_by(experiment, ID, goal_type) %>%
  complete(move_type = c("Useful", "Harmful", "Inconsequential"), fill = list(move_count = 0)) %>%
  mutate(total_moves = sum(move_count),
         prop = move_count / total_moves) %>%
  ungroup() %>%
  mutate(across(c(experiment, goal_type, move_type, ID), as.factor)) %>%
  select(-move_count, -total_moves)


prop_model = lm(data = prop_moves, prop ~ experiment*move_type*goal_type)
car::Anova(prop_model)

# Get EMMs from full model, averaged over goal_type
emm <- emmeans(prop_model, ~ experiment | move_type)

# Pairwise comparisons between experiments within each move_type
contrast_results <- contrast(emm, method = "pairwise", adjust = "tukey")

contrast_results
```
### plot
```{r}

ggplot(prop_moves, aes(x = goal_type, y = prop, fill = experiment)) +
  stat_summary(fun = mean, geom = "bar", position = position_dodge(width = 0.8), width = 0.7) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               position = position_dodge(width = 0.8), 
               width = 0.2) +
  facet_wrap(~ move_type, scales = "free_y") +
  labs(y = "Mean Proportion of Moves", x = "Goal Type") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal()
```


``` {r}


```



```{r}
```


# Follow-up measure of pragmatism seen through Helper's move utility following rank of Architect's first move (pragmatism rank). 

Found no significant correlation between architect first move rank and helper utility. 

```{r}
library(tidyverse)

# select architects' first‐move rank
arch <- read_csv("data/e1/final_first_moves.csv") %>%
  rename(id = ID) %>%
  select(id,goal, first_move_rank)

# select helper's first move utility
helpers <- read_csv("data/e1/final_tracker.csv") %>%
  rename(id = ID) %>%
  filter(role == "helper", move_id == 1) %>%
  select(id, goal, helper_move = move, helper_utility = move_utility) %>%
  mutate(
    helper_action = if_else(helper_move == "pass", "pass", "move")
  )

# join architect ranks to helper utilities
df_comb <- arch %>%
  inner_join(helpers, by = c("id","goal"))

# classify into useful, harmful etc.

df_comb %>% group_by(helper_action) %>% 
  summarize(mean_rank = mean(first_move_rank, na.rm = TRUE))%>%
  ggplot(aes(x = helper_action, y = mean_rank)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7)

# how often will helpers pass or move on their first move? 
df_comb %>%
  count(helper_action) %>%
  mutate(prop = n / sum(n))

# do more pragmatic arch moves reduce the chance of a pass?

# create a arch_rank column based on the median of first_move_rank, splitting into high and low rank
df_comb <- df_comb %>%
  mutate(arch_rank = if_else(first_move_rank <= median(first_move_rank, na.rm = TRUE), "high_prag", "low_prag"))

helper_pass_summary <- df_comb %>%
# count how many helper_actions occurred in each arch_rank
  count(arch_rank, helper_action) %>%
  group_by(arch_rank) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()


print(helper_pass_summary)



# Spearman correlation between first_move_rank and helper_utility

library(broom)
library(dplyr)

tidy(cor_result <- cor.test(df_comb$first_move_rank,
                    df_comb$helper_utility,
                    method = "spearman")) %>%
  transmute(
    rho = round(estimate, 2),
    p_value = round(p.value, 3),
  )



# visualize data with scatterplot
ggplot(df_comb, aes(x = first_move_rank, y = helper_utility)) +
  geom_jitter(width = 0.2, height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE, color = "steelblue") +
  labs(
    title = "Helper Utility vs. Architect First Move Rank",
    x = "Architect First Move Rank (lower = more pragmatic)",
    y = "Helper Move Utility"
  ) +
  theme_minimal()

# boxplot for group comparision
df_comb <- df_comb %>%
  mutate(
    arch_rank = factor(
      arch_rank,
      levels = c("high_prag", "low_prag"),
      labels = c("High Pragmatism", "Low Pragmatism")
  )
)
ggplot(df_comb, aes(x = arch_rank, y = helper_utility, fill = arch_rank)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, height = 0.05, alpha = 0.5, color = "black", size = 0.8) +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Helper Utility by Architect Pragmatism",
    x = "Architect Pragmatism",
    y = "Helper Move Utility"
  ) +
  theme_minimal()

```

# Extended to all moves

```{r}
library(tidyverse)
library(rstatix)
library(ggpubr)
library(broom)

# 1) Read & join architect ranks + helper first-move data
arch <- read_csv("data/e1/final_first_moves.csv") %>%
  rename(id = ID) %>%
  select(id, goal, first_move_rank)

helpers <- read_csv("data/e1/final_tracker.csv") %>%
  rename(id = ID) %>%
  filter(role == "helper", move_id == 1) %>%
  select(id, goal, helper_move = move, helper_utility = move_utility)

df_comb <- arch %>%
  inner_join(helpers, by = c("id","goal"))

# 2) Make a 4-level helper_class
df_comb <- df_comb %>%
  mutate(
    helper_class = case_when(
      helper_move == "pass"         ~ "pass",
      helper_utility ==  1          ~ "useful_move",
      helper_utility ==  0          ~ "inconsequential_move",
      helper_utility == -1          ~ "harmful_move"
    ) %>%
    factor(levels = c("pass","useful_move","inconsequential_move","harmful_move"))
  )

# 3) Frequencies & proportions of each class
df_comb %>%
  count(helper_class) %>%
  mutate(prop = n / sum(n)) %>%
  print()

# 4) Mean ± SD of architect first_move_rank by helper_class
df_comb %>%
  group_by(helper_class) %>%
  summarise(
    mean_rank = mean(first_move_rank, na.rm = TRUE),
    sd_rank   = sd(first_move_rank,   na.rm = TRUE),
    n         = n(),
    .groups   = "drop"
  ) %>%
  print()

```




```{r}


```




